\chapter{Conclusions \& Future Work}
\label{chapt:conclusions}

Through a combination of differencing and proof term transformations,
my proof repair tool suite can extract, generalize, and apply the information that a change carries to fix proofs broken by the same change.
Proof repair can save and in fact already has saved work for proof engineers relative to reference manual repairs in practical use cases.
And so proof repair is \kl{reason to believe} that verifying a modified system should often, in practical use cases, be easier than verifying the original the first time around,
even when proof engineers do not follow good development processes,
or when change occurs outside of proof engineers' control.

Do not just take my word for it.
Consider a recent article by a proof engineer saying just this (emphasis mine): % TODO link: https://galois.com/blog/2020/12/proofs-should-repair-themselves/

\begin{quote}
We have \textit{reason to think} such proof repair is tractable. Rather than trying to synthesize a complete proof from nothing---a problem known to be immensely difficult---we 
start from a correct proof of fairly similar software. We will be attempting proof reconstruction \textit{within a known neighborhood}.
\end{quote}
The proof engineer credited my proof repair work on Twitter, % TODO link to Twitter
but noted that there ought to be much more work in this space.

I agree.
Actually, I want to take that a step further:
proof repair is just a small piece of what will carry us to the next era of proof engineering.
And that era will be one in which programmers of all skill levels across all domains can develop and maintain verified systems---the era of
\textit{proof engineering for all}.

\input{next}
