\chapter{Motivating Proof Repair}
\label{chapt:mot}

This thesis describes techniques and tools for automatically \kl{repairing} broken \kl{proofs} in a \kl{proof assistant}.
It focuses in particular on proofs about formally \kl{verified} \kl{programs}, though many of the techniques and tools carry over to mathematical proofs as well.

\paragraph{Why Verify Programs?}

Formal verification of a program can improve actual and perceived reliability.
It can help the programmer think about the desired and actual behavior of the program,
perhaps finding and fixing bugs in the process~\cite{murraybp}.
It can make explicit which parts of the system are trusted, and further decrease the burden
of trust as more of the system is verified.

One noteworthy program verification success story is the CompCert~\cite{Leroy:POPL06, Leroy2009} verified optimizing C compiler.
Both the back-end and front-end compilation passes
of CompCert have been verified, ensuring the correctness of their composition~\cite{Kaestner2017}.
CompCert has stood up to the trials of human trust: it has been used, for example, to compile code for safety-critical flight control software~\cite{Frana2011}.
It has also stood up to rigorous testing: while the test generation tool Csmith~\cite{Yang2011} found 
79 bugs in GCC and 202 bugs in LLVM, it was unable to find any bugs in the verified parts of CompCert.
%This demonstrated the empirical reliability of CompCert, which in turn further increased confidence in its reliability.

CompCert, however, was not a simple endeavor: the original development comprised approximately 35000 lines of code;
functionality accounted for only 13\% of this, while specifications and proofs accounted for the other 87\%.
This is not unusual for large proof developments. The initial correctness proofs for the seL4 OS microkernel,
for example, consisted of 480000 lines of specifications and proofs~\cite{Klein2014micro}.
\kl{Proof engineering} technologies make it possible to develop verified systems at this scale.
See \kl{QED at Large} for a comprehensive overview of proof engineering.

\paragraph{Why Repair Proofs?}

\kl{Proof repair}---the focus of this thesis---is a new proof engineering technology that focuses in particular on minimizing the burden of change as verified systems evolve over time.
But for the sake of this chapter, I motivate proof repair not on a large verified system like a C compiler or an OS microkernel.
Instead, I motivate it on a simple proof development: a list zip function accompanied by a formal proof that it preserves the lengths of its inputs.
This is a small example, but it is worth noting that large proof developments like compilers and OS microkernels
are often made up of many of these small examples built on top of each other.

The proof assistant that I motivate this on is \kl{Coq}, since this is the proof assistant that this thesis focuses on.
I motivate proof repair in Coq using the small list zip example in three parts:

\begin{enumerate}
\item the workflow of and theory beneath \textit{proof development} (Section~\ref{sec:mot-dev}),
\item some challenges of and approaches to \textit{proof maintenance} (Section~\ref{sec:mot-mai}), and
\item the motivation for and approach to \textit{proof repair} that follow (Section~\ref{sec:mot-rep}).
\end{enumerate}
I will refer to the example and theory introduced in this chapter in later chapters, so it is good to at least skim this chapter regardless of Coq experience.

%I think Chapter 2 probably has the right content you need for what follows.  The order /within/ Chapter 2 probably makes it more challenging to readers (or at least me) than is needed.  Would it work to reorder and refactor something like this:
%  * Current 2.0 and 2.1.0
%  *  But then 2.1.1.2 List Zip Preserves Length as a 2.X -- just for the moment take, as people actually do in practice, the standard library's types and functions as 'primitives'
%  * *Then* say, actually, those aren't built-in to Coq, they're just inductive types/definitions and we will take them as defined in terms of eliminators not pattern matching and such.
%  * *Then* go even lower-level and say, actually this is all CIC

%Okay, so now for proof maintenance, we can see what's going on at all three of the previous levels: what happens to programmers, why the proofs break, what this means at the most foundational level

%And then what proof repair can do about it.

\input{development}

\input{maintenance}

\input{repair}
